\chapter{Intellectual Property}

\section{Intellectual Property Protection}
\subsection{Copyright}
Copyright gives a creator the right to control how their creative work may be used. It applies to a work, rather than the concept behind it.

It is automatic so you do not have to apply or pay a fee. The copyright symbol is not needed except as a reminder. Employers may have contracts (a `contract of employment') that gives them copyright over work produced by their employees.

The copyright will last until 70 years after the creator's death. Copyrighted works can be used without permission for:
\begin{itemize}
    \item Non-commercial research
    \item Criticism, review, and reporting current events
    \item Teaching
    \item Helping disabled people for instance by changing the format of a book
    \item Parody, caricature, and pastiche
\end{itemize}
Software is easy to make a perfect copy of, unlike most other works.

\subsection{Copyleft}
Richard Stallman created a list of freedoms for free software:
\begin{itemize}
    \item To run the program for any purpose
    \item To access the source code
    \item To redistribute exact copies
    \item To redistribute with changes
\end{itemize}
In 2001, Microsoft CEO Steve Ballmer said: `Linux is a cancer that attaches itself in an intellectual-property sense to everything it touches.' In 2018 the new CEO Satya Nadella bought GitHub for \$7.5 Billion.

OpenSSL has only two full time employees to maintain, test, and review 500,000 lines of business critical code.

Various companies have bug bounty programs that allow anyone to get paid for finding bugs in their software.

Open source software normally has a licence to control its use.

\begin{example}
    Should search engines compensate copyright owners for storing and displaying cached versions of their websites?

    In \emph{Field v Google} a court ruled that it is fair use for Google to display cached versions of a copyrighted website.
\end{example}

\subsection{Trademarks}
A trademark is an intellectual property that enables one to identify goods or services from a specific source. Registering trademarks is not automatic.

A trademark cannot be:
\begin{enumerate}
    \item Offensive
    \item Descriptive
    \item Misleading
    \item Three dimensional
    \item Non-distinctive
    \item Used
\end{enumerate}
Cybersquatting is the action of using an internet domain name with intent to profit from a trademark belonging to someone else. A similar process is done with social media usernames.

\subsection{Patents}
For a patent to be issued, the underlying must be new, involve an inventive step, and be capable of industrial application. The patent will then last for 20 years.

Companies like \emph{Intellectual Ventures} buy large numbers of patents and make money from royalties for their use. 90\% of these cases never go to court, as settlements are cheaper than going to court.

Patents have the disadvantage that in order to file for one any trade secrets will be lost. This can be avoided using a Non-disclosure agreement (NDA), that can be unilateral, bilateral, and multilateral.
